---
title: "NBA Draft Markdown"
author: "Mohammed Khan"
date: '2022-06-16'
output: html_document
---

# TO-DO

## clean data for 18-21 
## train model with data 09-21
## run model on 22
## clean ggplot
## check training accuracy by comparing max prob with cluster class

# P1 |> 
## plot PCA
### basic v added condor
## generate table
## show outliers
## by draft position

# RESOURCES

## Regression
### tutorial : https://stats.oarc.ucla.edu/r/dae/multinomial-logistic-regression/

## Pace and Space 
### trends : https://trends.google.com/trends/explore?cat=264&date=all&geo=US&q=pace%20and%20space

## Predicted Probabilities Notes 09-17
### Marcus Smart
## Nerlens Noel
## Kawhi Leonard
### 2.0+ college DWS / average condor ~7 inches 

## Meyers Leonard
## Willie Warren
## Lorenzo Brown


```{r setup}
library(tidyverse) # data manipulation
library(factoextra) # clustering viz
library(data.table) # for fread
library(progress)
# load progress bar / call pb$tick(0)
pb <- progress_bar$new(total = 100)
# recent db
height_weight <- read.csv("measurements.csv")
# draft db
dl20 <- read.csv("2020-2020 NBA Draft Lottery.csv")
dl21 <- read.csv("2021-2021 NBA Draft Lottery.csv")
# for missing wingspans
wing2018 <- read.csv("2018 wingspan.csv")
wing2019 <- read.csv("2019 wingspan.csv")
wing2020 <- read.csv("2020 wingspan.csv")
wing2021 <- read.csv("2021 wingspan.csv")
# 2009 - 2017
combine <- read.csv("combine_all.csv")
```

```{r clean}
# one entry per player
hw_test <- height_weight |> 
  distinct(Player, .keep_all = TRUE)
# convert feet-inches to inches for hw 
hw_test.1 <- hw_test |> 
  janitor::clean_names() |> 
  separate(height, c('feet', 'inches'), sep = '-', convert = TRUE, remove = FALSE) |> 
  mutate(height = 12 * feet + inches) |> 
  select(-c(feet, inches)) |> 
  filter(draft_year >= 2018 & draft_year != c("Undrafted", ""))
# remove height with shoes 
combine.1 <- combine |> 
  select(-Height..With.Shoes.) |> 
  rename(Height = Height..No.Shoes.) |> 
  janitor::clean_names() |> 
  mutate(condor = wingspan - height)
dl20.test <- dl20 |> 
  janitor::clean_names() |> 
  select(last, height, wingspan, weight) |> 
  mutate(condor = wingspan - height)
dl21.test <- dl21 |> 
  janitor::clean_names() |> 
  select(last, height, wingspan, weight) |> 
  mutate(condor = wingspan - height)
```

```{r explore}
condor_metric <- combine.1 |> 
  select(condor, bench, agility, sprint) |> 
  group_by(condor) |> 
  summarise(across(everything(), .f = list(mean = mean), na.rm = TRUE))
```

```{r cluster}
# prep for clustering 
combine.clus.basic <- combine.1 |> 
  select(height, wingspan, weight) |> 
  na.omit()
combine.clus.all <- combine.1 |> 
  select(height, wingspan, weight, condor) |> 
  na.omit()
test <- combine.1 |> 
  select(player, height, wingspan, weight, condor, bench, agility, sprint) |> 
  na.omit()
# kmeans 
set.seed(1234)
kmeans.basic <- kmeans(x = combine.clus.basic, centers = 5, nstart = 25)
kmeans.all <- kmeans(x = combine.clus.all, centers = 5, nstart = 25)
# attach clusters to data 
combine.clus.basic$cluster <- kmeans.basic$cluster
combine.clus.all$cluster <- kmeans.all$cluster
#combine.1$cluster <- as.factor(combine.clus.basic$cluster)
# viz kmeans 
pca.basic <- fviz_cluster(kmeans.basic, combine.clus.basic)
pca.all <- fviz_cluster(kmeans.all, combine.clus.all)
pca.basic
pca.all
combine.clus.basic$player <- combine.1$player
combine.clus.all$player <- combine.1$player
# recode clusters to positions 1-5 for basic
combine.basic <- combine.clus.basic |> 
  mutate(position = recode(cluster, "1" = 1, "2" = 3, "3" = 5 ,"4" = 2, "5" = 4)) |> 
  mutate(position = as.factor(position)) |> 
  select(c(player, height, wingspan, weight, position))
# recode clusters to positions 1-5 for all
combine.all <- combine.clus.all |> 
  mutate(position = recode(cluster, "1" = 1, "2" = 4, "3" = 5, "4" = 3, "5" = 2)) |> 
  mutate(position = as.factor(position)) |> 
  select(c(player, height, wingspan, weight, condor, position))
# means by position check for basic
combine.mean.basic <- combine.basic |> 
  select(-player) |> 
  group_by(position) |> 
  summarise(across(everything(), mean)) |> 
  # recode default cluster numbers to postion numbers 
  ## possible change to if/else 
  mutate(position = recode(position, 
                           "1" = "1's", "2" = "2's", "3" = "3's", "4" = "4's", "5" = "5's")) |> 
  # change case of columns to capital first letter, rest lowercase
  rename_with(str_to_title) |> 
  # create condor variable from height and weight 
  mutate(Condor = Wingspan - Height) |> 
  # round all numerical values to 2 decimal places 
  mutate_if(is.numeric, round, digits = 1) |> 
  relocate("Condor", .after = "Wingspan")
# means by position check for all
combine.mean.all <- combine.all |> 
  select(-player) |> 
  group_by(position) |> 
  summarise(across(everything(), mean))
# means by pace and space
```

```{r gt_table}
require(gt)
require(gtExtras)
combine.mean.basic |> 
  gt() |> 
  gt_theme_538() |> 
  tab_header(title = "Average Measurements of Height (in.), Wingspan (in.), 
             and Weight (lbs.) By Position",
             subtitle = "Note: Condor is the difference between Wingspan and Height") |> 
  gt::data_color(
    columns = Height:Weight, colors = c("Red", "Black")
  )
```


```{r ggplot}
pca.basic + 
  ggtitle("NBA Player Clusters \nby Height, Wingspan, and Weight") + 
  scale_fill_discrete(name = "Position", labels = c("Ones", "Threes", "Fives", "Twos", "Fours")) + 
  theme_classic()
  
```

```{r multinomial regression basic}
require(foreign)
require(nnet)
require(ggplot2)
require(reshape2)

combine.basic$position_2 <- relevel(combine.basic$position, ref = "1")
train.basic <- multinom(position_2 ~ height + wingspan + weight, 
                        data = combine.basic)
summary(train.basic)
z <- summary(train.basic)$coefficients / summary(train.basic)$standard.errors
z
p <- (1 - pnorm(abs(z), 0, 1)) * 2
p
exp(coef(train.basic))
# predicted probabilities of position
head(pp <- fitted(train.basic))
combine.basic$pp <- round(pp, 4)
```

```{r multinomial regression all}
require(foreign)
require(nnet)
require(ggplot2)
require(reshape2)

combine.all$position_2 <- relevel(combine.basic$position, ref = "1")
train.all <- multinom(position_2 ~ height + wingspan + weight + condor, 
                        data = combine.all)
summary(train.all)
z <- summary(train.all)$coefficients / summary(train.all)$standard.errors
z
p <- (1 - pnorm(abs(z), 0, 1)) * 2
p
exp(coef(train.all))
# predicted probabilities of position
head(pp <- fitted(train.all))
combine.basic$pp <- round(pp, 4)
```

```{r test}
dl20.test$pp.basic1 <- predict(train.basic, newdata = dl20.test, type = "class")
dl20.test$pp.basic2 <- round(predict(train.basic, newdata = dl20.test, type = "prob"), 4)

dl21.test$pp.basic1 <- predict(train.basic, newdata = dl21.test, type = "class")
dl21.test$pp.basic2 <- round(predict(train.basic, newdata = dl21.test, type = "prob"), 4)

dl20.test$pp.all1 <- predict(train.all, newdata = dl20.test, type = "class")
#dl20.test$pp.basic2 <- round(predict(train.all, newdata = dl20.test, type = "prob"), 4)

dl21.test$pp.all1 <- predict(train.all, newdata = dl21.test, type = "class")
#dl21.test$pp.basic2 <- round(predict(train.all, newdata = dl21.test, type = "prob"), 4)
```





